---
title: "Covid91 vaccine study Final2021"
author: "Adriana Pimentel"
date: "11/23/2021"
output: 
    html_document:
    toc: true
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

FinalData <-FinalData %>% mutate_if(is.character,as.factor)
FinalData$treatment <- relevel(FinalData$treatment, ref = "placebo")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Guys

```{r}
Guys <- filter(FinalData, sex =="guy")
```

```{r}
barchartGC(~infected + treatment,data=Guys)
barchartGC(~infected + treatment,data=Guys, type="percent")
```

The bar charts show that the majority of the guys population were not infected. However, the guys who received the drug were more likely to get infected than the guys who received the placebo. 

```{r}
table1 <- xtabs(~infected + treatment, data=Guys)
rowPerc(table1)
colPerc(table1)
```

Here the table shows that about 58.2 percent of the guys who took the drug had covid and about 41.8 percent of the guys who took the placebo had covid. Proving that, there were more infected guys who took the treatment versus the ones who took the placebo. There were more guys uninfected when taking the placebo because there was about 51.5 percent guys with the placebo who were ok. 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Guymod <- fisher.test(table1)
guyeff <- (Guymod$estimate/(Guymod$estimate + 1))*100
guyeffupper <-Guymod$conf.int[2] 
guyeffuppereff<- (Guymod$conf.int[2]/(Guymod$conf.int[2]+1))*100
guyefflower <-Guymod$conf.int[1]
guyefflowereff <-(Guymod$conf.int[1]/(Guymod$conf.int[1]+1))*100
options(digits=3)
```
  
The given p-value for the chi square test and the Fisher's Exact test are both low. Indicating that the drug taken does have a significant affect on whether or not the guys get infected with COVID. The Fisher's test calculated that the odds of getting COVID when taking the drug is `r Guymod$estimate`. Which is based on the 95% confidence interval ranging from `r Guymod$conf.int[1]  ` to `r Guymod$conf.int[2] ` . Therefore, we reject the null hypothesis according to the tests shown above.I would not recommend the treatment.   
  
The drug is only `r guyeff`% effective on Guys, which means it is not effective. The confidence interval is from `r guyefflowereff ` to `r guyeffuppereff`. According to these results, I would not recommend Guys to receive the covid-19 study drug.

# Gals



# LGBTQ

# Drug Users

# Discussion/Conclusion

